---
title: "p8105_hw3_yh3676"
author: "Yuzhe Hu"
date: "2023-10-13"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 9, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```
 
#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

#### Read in `brfss_smart2010` data

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

#### Do data cleaning

```{r}
brfss_df = 
  brfss_smart2010 |>
  # format the data to use appropriate variable names
  janitor::clean_names() |> 
  rename(state = locationabbr, location = locationdesc) |>
  # focus on the "Overall Health” topic
  filter(topic == "Overall Health") |>
  # include only responses from “Excellent” to “Poor”
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor" )) |>
  # organize responses as a factor taking levels ordered from “Poor” to “Excellent”
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

#### Using this dataset, answer the following three questions
In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
table_2002 =
  brfss_df |>
  filter(year == 2002) |>
  group_by(state) |> 
  summarise(location_count = n_distinct(location)) |> 
  filter(location_count >= 7)

table_2010 =
  brfss_df |>
  filter(year == 2010) |>
  group_by(state) |> 
  summarise(location_count = n_distinct(location)) |> 
  filter(location_count >= 7)
```
* In 2002, `r table_2002 |> pull(state)` were observed at 7 or more locations.
* In 2010, `r table_2010 |> pull(state)` were observed at 7 or more locations.

Construct a dataset that is limited to `Excellent` responses, and contains, year, state, and a variable that averages the `data_value` across locations within a state. Make a “spaghetti” plot of this average value over time within a state.

```{r, message=FALSE}
# Construct the dataset
excellent_df =
  brfss_df |>
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarise(data_value_mean = mean(data_value, na.rm = TRUE))

# Make a `spaghetti` plot
excellent_df |>
  ggplot(aes(x = year, y = data_value_mean, group = state, color = state)) +
  geom_line() +
  labs(title = "Spaghetti Plot of Average Value Over Time within each State",
       x = "Year",
       y = "Average Value") +
theme(legend.position = "right")
```

* It can be seen that the average data value varies among the years as well as different states.

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_df |>
  filter(state == "NY" , year %in% c(2006, 2010)) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(title = "Distribution of data_value by Response in NY State",
       x = "response",
       y = "data_value") +
  facet_grid(. ~ year) +
  theme_bw()
```

* It can be seen that there is a significant increase in the `data_value` for `Very Good` response in NY State from 2006 to 2010.


## Problem 3

#### Load, tidy, merge, and otherwise organize the data sets

Final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r covar, message=FALSE}
covar_df = 
  read_csv('data/nhanes_covar.csv', skip = 4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
  drop_na() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ 'male',
      2 ~ 'female')) |>
  mutate(
    education = case_match(
      education,
      1 ~ 'Less than high school',
      2 ~ 'High school equivalent',
      3 ~ 'More than high school')) |>
  mutate(education = factor(education, levels = c('Less than high school', 'High school equivalent', 'More than high school'), ordered = TRUE))

accel_df =
  read_csv('data/nhanes_accel.csv') |>
  janitor::clean_names() 

joined_df = 
  left_join(covar_df, accel_df, by = 'seqn') 
```

#### Produce a reader-friendly table for the number of men and women in each education category

```{r, message=FALSE}
joined_df |> 
  group_by(sex,education) |> 
  summarise(number = n()) |> 
  pivot_wider(
    names_from = education, 
    values_from = number
  ) |>
  knitr::kable()
```

* From the table, it can be seen that the participants’ sex-ratio are mostly balanced in `Less than high school` category and `More than high school` category. In `High school equivalent `category, there are more male participants than female (35:23). The `More than high school education` category contains more participants compared to the other two categories.


#### Create a visualization of the age distributions for men and women in each education category

```{r}
joined_df |>
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distribution by Education and Gender", x = "Education", y = "Age")
```

* From the boxplot, it can be seen that participants with `More than high school` education is significantly younger than the other two education categories. In `High school equivalent` category, male participants are younger than female participants, while in the other two categories the age distributions are similar between sexes.


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.
